Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-26 robdml.tex
tex

Even if $H_\mathsf{D}$ is calucated from another triplet harder than the
\emph{source} triplet, the adversary may become weak in the late phase of
training.
%
The training process of the model aims to reduce the expectation of triplet loss
$E[L_\text{T}]$ to zero over the distribution of sample triplets,
and thus the $E[H]$ of any given triplet will eventually tend to $-\gamma$,
reducing the hardness of adversarial triplets from HM.

\cref{fig:ga}
% hardness boosting

Such deficiency in the late training process can be alleviated with a
pseudo-hardness function that boosts the destination hardness when the loss
value is low.
%
Denoting the loss value from the previous training iteration as $\ell_{t-1}$,
we first normalize it into $[0,1]$ as $\bar{\ell}_{t-1}=\min(u,\ell_{t-1})/u$,
where $u$ is a manually specified parameter.
%
Note, $u$ is much less than the theoretical upperbound of the loss function,
in order to avoid excessive hardness boost in the functions introduced below.
%
Then the expectation of destination hardness can be linearly shifted (boosted)
by a scaled constant by a ``hardness boosting'' pseudo-hardness
function, \ie,
%
\begin{equation}
	%
	g_\mathsf{B}(H_\mathsf{D};\xi,\bar{\ell}_{t-1}) =
	H_\mathsf{D} + \xi \cdot (1-\bar{\ell}_{t-1}).
	%
\end{equation}
%
And the adversarial examples from $\text{HM}(H_\mathsf{S},g_\mathsf{B}(H_\mathsf{D}))$
can be used for training.

@@

>>>>

Even if $H_\mathsf{D}$ is calucated from another triplet harder than the
\emph{source} triplet, the adversarial example may become weak in the late
phase of training.
%
The optimizer aims to reduce the expectation of triplet
loss $E[L_\text{T}]$ to zero as possible over the distribution of sample triplets, and thus
the $E[H]$ of any given triplet will tend to $-\gamma$, reducing the
hardness of adversarial triplets from HM.
%
We speculate this will lower the level of robustness the model could reach.

% hardness boosting

Intuitively, such deficiency can be alleviated with a \emph{pseudo-hardness}
function that slightly increase the value of $H_\mathsf{D}$ in the late phase
of training.
%
Denoting the loss value from the previous training iteration as $\ell_{t-1}$,
we first normalize it into $[0,1]$ as $\bar{\ell}_{t-1}=\min(u,\ell_{t-1})/u$,
where $u$ is a manually specified constant.
%
Then we can linearly shift the $E[H_\mathsf{D}]$ by a scaled constant $\xi$,
\ie,
%
\begin{equation}
	%
	g_\mathsf{B}(H_\mathsf{D};\xi,\bar{\ell}_{t-1}) =
	H_\mathsf{D} + \xi \cdot (1-\bar{\ell}_{t-1}).
	%
\end{equation}
%
The deficiency can be alleviated in
$\text{HM}[H_\mathsf{S},g_\mathsf{B}(H_\mathsf{D})]$.

\cref{fig:ga}

@@

Note, $u$ is much less than the theoretical upperbound of the loss function,
in order to avoid excessive hardness boost in the functions introduced below.
